Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-05-26 2017-Xudong-SMT/AVOCS-2018-pre-proceeding.tex
Hide the appendix in pre-proceeding.
Improve the pre-proceeding paper.

For other applications, such as the analysis of 
programming languages, there exist dedicated platforms using
external automatic theorem 
provers (ATP) or automatic tactics from interactive theorem
provers (ITP), to perform symbolic reasoning, and for example to
discharge some subgoals in the proofs.
Tools like Rodin~\cite{deharbe2013,deharbe2014,abrial2007} have
already integrated several provers, like Z3, as modules for proving
the proof obligations generated from a user model. 
The prover we use, which also happens to be Z3, is developed by Microsoft Research
based on the satisfiability modulo 
theories framework (SMT), is mainly applied in extended static checking, test case
generation, and predicate abstraction.
In a similar way, there are several ATPs/ITPs we could consider to use for
result pruning and bisimulation checking in our algorithm, as an
alternative to Z3, such as CVC4~\cite{barrett:CAV2011},
Coq~\cite{armand:CPP2011},
Isabelle~\cite{blanchette:FroCoS2011} or others.

@@

\begin{example}[An open transition]
  This transition is generated by application of the vector \texttt{SV0},
  synchronizing the initial actions of pLTSs \texttt{Control} and
  \texttt{Timer} (see Fig.~\ref{schema:ArchFailure:pNet}), with an action of the hole \texttt{B} equal to
  \texttt{fail}. The local variable \texttt{t} of the \texttt{Timer}
  is assigned to its initial value \texttt{Max}.
 A full output of 
the use-case is provided in ~\cite{RR}.

@@

\paragraph{Structural Semantic Rules:}
The semantics of pNet in term of open automata has been defined in
\cite{henrio:Forte2016}, in the form of 2 structural rules, one for
pLTSs, one for pNet nodes. These rules are improved, adding guards in the synchronisation vectors and quantifier in the guards, in ~\cite{RR}.

@@

The inside loop (\emph{MakeTransitions} method) applies recursively
the semantic rules following the structure of the pNet.
When applied to a pLTSs at the leaves, we simply take the pLTS transitions of the
corresponding local state and use the semantic rule to build the OT.\footnote{%
%
  We omit detailed presentation of this case for the sake of brevity.
%
}
When applied to a pNet node we use two methods, combining and
matching, to generate the open transitions in a hierarchical manner,
as shown in Alg.~\ref{alg2}. This method directly manages the
holes of the node, so \emph{MakeTransitions} is never called on a hole.

@@

Our implementation submits satisfiability requests to
Z3 using its JAVA API. Here, for readability, we
show the Z3 code using its SMT-LIB input language.
Note that in the previous sections, the OTs were displayed in a
  simplified, human readable form. The input and output of our tool, and
  also the generated SMTlib fragments, are
  slightly more difficult to read, in 
particular because of structured names for the fresh variables generated
by the algorithm, allowing tracability of the result~\cite{RR}.

@@

\emph{Translation of action algebra presentation.}
In~\cite{RR}, we  define the condition of
well-formedness of an algebra presentation, and its translation into
SMTlib declarations (\texttt{declare-datatypes} and
\texttt{declare-fun}). It ensures that the generated code is correct,
and will not raise runtime errors in the SMT engine. Note that the
\texttt{declare-datatypes} command comprises both the action
constructors from table \ref{table:BIPalgebra} and also the constant action
names from our example.
Additionally, we will include axiomatisation of any required functions
and predicates of the presentation data-types.

@@

\emph{Translation of open transitions.}
In~\cite{RR} we formally define all steps of the translation of each
open transition, including:
%
\begin{itemize}
\item collect all variables in the transition, and declare them (using \texttt{declare-const})
\item check well-formedness and correct typing of expressions,
\item translate the predicate into a conjunction of assertions,
\item if present in the source state, translate the state-variable
  assignments into a disjunctive assertion.
\end{itemize}
%
This translation ensures that no runtime error will occur
in the SMT engine.

@@

\emph{Scaling up.} One important motivation of this work is to
attack the complexity of verification of realistic systems by a
compositional and parametric approach. Still one may wonder if the price
for analysing our symbolic transitions will not make the approach too expensive
in term of computing time. We tried a
slighly bigger example, assembling 2 Failure controllers. In~\cite{RR}, we show that Z3 can check the
satisfiablility of a 90K open transitions in a couple of minutes.

@@

\newpage
\appendix
%\section{Appendix}
The following appendices are included for the convenience of reviewers.
Our intention is to publish as a research report, before final publication, an
extended version containing all technical details, proofs, and more examples.

>>>>

For other applications, such as the analysis of 
programming languages, there exist dedicated platforms using
external automatic theorem 
provers (ATP) or automatic tactics from interactive theorem
provers (ITP), to perform symbolic reasoning, and for example to
discharge some subgoals in the proofs.
Tools like Rodin~\cite{deharbe2013,deharbe2014} have
already integrated several provers, like Z3, as modules for proving
the proof obligations generated from a user model. 
The prover we use, which also happens to be Z3, is developed by Microsoft Research
based on the satisfiability modulo 
theories framework (SMT), is mainly applied in extended static checking, test case
generation, and predicate abstraction.
In a similar way, there are several ATPs/ITPs we could consider to use for
result pruning and bisimulation checking in our algorithm, as an
alternative to Z3, such as CVC4~\cite{barrett:CAV2011},
Coq~\cite{armand:CPP2011} or others.

@@

\begin{example}[An open transition]
  This transition is generated by application of the vector \texttt{SV0},
  synchronizing the initial actions of pLTSs \texttt{Control} and
  \texttt{Timer} (see Fig.~\ref{schema:ArchFailure:pNet}), with an action of the hole \texttt{B} equal to
  \texttt{fail}. The local variable \texttt{t} of the \texttt{Timer}
  is assigned to its initial value \texttt{Max}.
 A full output of 
the use-case is provided in~\cite{Avocs-RR}.

@@

\paragraph{Structural Semantic Rules:}
The semantics of pNet in term of open automata has been defined in
\cite{henrio:Forte2016}, in the form of 2 structural rules, one for
pLTSs, one for pNet nodes. These rules are slightly improved, adding guards in the synchronisation vectors and syntax for universal quantifier in the guards (see~\cite{Avocs-RR}).

@@

The inside loop (\emph{MakeTransitions} method) applies recursively
the semantic rules following the structure of the pNet.
When applied to a pLTSs at the leaves, we simply take the pLTS transitions of the
corresponding local state and use the semantic rule to build the OT\footnote{%
%
  We omit detailed presentation of this case for the sake of brevity.
%
}.
When applied to a pNet node we use two methods, combining and
matching, to generate the open transitions in a hierarchical manner,
as shown in Alg.~\ref{alg2}. This method directly manages the
holes of the node, so \emph{MakeTransitions} is never called on a hole.

@@

Our implementation submits satisfiability requests to
Z3 using its JAVA API. Here, for readability, we
show the Z3 code using its SMT-LIB input language.
Note that in the previous sections, the OTs were displayed in a
  simplified, human readable form. The input and output of our tool, and
  also the generated SMTlib fragments, are
  slightly more difficult to read, in 
particular because of structured names for the fresh variables generated
by the algorithm, allowing tracability of the result~\cite{Avocs-RR}.

@@

\emph{Translation of action algebra presentation.}
In~\cite{Avocs-RR}, we  define the translation of an algebra presentation into
SMTlib declarations (\texttt{declare-datatypes} and
\texttt{declare-fun}). We also formalize a condition of well-formedness of pNets ensures that the generated code is correct,
and will not raise runtime errors in the SMT engine. Note that the
\texttt{declare-datatypes} command comprises both the action
constructors from table \ref{table:BIPalgebra} and also the constant action
names from our example.
Additionally, we will include axiomatisation of any required functions
and predicates of the presentation data-types.

@@

\emph{Translation of open transitions.}
In~\cite{Avocs-RR} we formally define all steps of the translation of each
open transition, including:
%
\begin{itemize}
\item collect all variables in the transition, and declare them (using \texttt{declare-const})
\item check well-formedness and correct typing of expressions,
\item translate the predicate into a conjunction of assertions,
\item if present in the source state, translate the state-variable
  assignments into a disjunctive assertion.
\end{itemize}
%
This translation ensures that no runtime error will occur
in the SMT engine.

@@

\emph{Scaling up.} One important motivation of this work is to
attack the complexity of verification of realistic systems by a
compositional and parametric approach. Still one may wonder if the price
for analysing our symbolic transitions will not make the approach too expensive
in term of computing time. We tried a
slighly bigger example, assembling 2 Failure controllers. In~\cite{Avocs-RR}, we show that Z3 can check the
satisfiablility of a 90K open transitions in a couple of minutes.

@@

%\newpage
%\appendix
%\section{Appendix}
The following appendices are included for the convenience of reviewers.
Our intention is to publish as a research report, before final publication, an
extended version containing all technical details, proofs, and more examples.
